---
title: 'Exploration of Health Data of Hugo Certuche'
author: "Hugo  Certuche hsc539"
date: ''
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = F, message = F,
tidy=TRUE, tidy.opts=list(width.cutoff=60), R.options=list(max.print=100))
```




*This project was a very interesting one and as such I thought loud and hard about what data I was going to use for it! I am quite passionate about health and fitness. The variables that  I collected using MyFitnessPal are Calories, which is the number that I consumed each day and EnoughFiber which is whether or not I got the recommended amount of fiber for the day. The variables that I got from my workout tracker are MaxHR, which is my maximum heart rate, AvgHR, which is my average heart rate, and WoDuration which is the duration of my workout. I expect that there will be a positive correlation between AvgHR and MaxHR.*


1 Datasets
```{R}
    
library(tidyverse)
HeartRateStats<-read_csv("HeartRateStatsss.csv")
HR<-HeartRateStats
HR
HeartRateStats
MyFitnessPalStatsProject <-read_csv("MyFitnessPalStatsProject.csv")
HR 
HR %>% pivot_longer(cols = c("Day","MaxHR","AvgHR","WoDuration"), names_to="varr",values_to="value")
```
*I loaded in my datasets for my workout stats and food stats. My data was already tidy but I did use pivot_longer to demonstrate how that I can manipulate the data for my workout data accordingly.*

2. Joining
```{R}
bigboi <-full_join(HR, MyFitnessPalStatsProject, by=c("Day"))



bigboi

```
*I did a full_join for this datasets and had no variables that were dropped from the dataset! I choose a full join because I did not want anything to be dropped.*

3. Summary Statistics
```{R}    
    
bigboi %>% mutate(pulses = AvgHR * WoDuration)
bigboi %>% filter(EnoughFiber=="Yes") %>% arrange(Calories)
bigboi %>% select(MaxHR,AvgHR,Calories)
bigboi %>% group_by(Day)  
    


bigboi %>% summarise_all(mean)

bigboi %>% summarise_all(sd)

bigboi %>% group_by(EnoughFiber) %>% summarize_all(mean)

bigboi %>% group_by(EnoughFiber) %>% summarize_all(sd)

bigboi %>%  summarize_all(n_distinct)

bigboi %>%  summarize_all(max)

bigboi %>%  summarize_all(min)


cormut <-bigboi %>% select_if(is.numeric) %>% cor(use="pair")



cormut

tidycorr<-cormut %>% as.data.frame %>% rownames_to_column("var1") %>% pivot_longer(-1,names_to="var2",values_to="correlation")

tidycorr

``` 
*I used all six dplyrs in various forms. I made a new variable called pulses that represents number of heart beats during a workout. I also got the means and standard deviations of all the variables. I also calculated the means and standard deviations for each variables after grouping by fiber content. Last I made a correlation matrix. I mainly noticed that in the groups of EnoughFiber, the enough fiber group had much more calories than the not enough fiber. The other thing I noticed is that group with more fiber also had a longer mean workout duration.*

4. Visualizations 
```{R}    
 tidycorr%>%ggplot(aes(var1,var2,fill=correlation))+  geom_tile()+  scale_fill_gradient2(low="red",mid="white",high="blue")+ geom_text(aes(label=round(correlation,2)),color = "black", size = 4)+  theme(axis.text.x = element_text(angle = 90, hjust=1))+ coord_fixed() +theme_gray()



ggplot(bigboi, aes(x = AvgHR, y = MaxHR, color=EnoughFiber, size=2 )) + theme(legend.position="none") +   ggtitle("MaxHR As A Function of AvgHR") + ylab("MaxHR") + xlab("AvgHR") + geom_point()+theme_dark() +scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
                                                                                                                                                                                      
ggplot(bigboi, aes(x= WoDuration, y=Calories, color="orange")) + geom_bar(stat = "summary", fun=mean)+geom_errorbar(stat = "summary", fun.data=mean_se) + ggtitle("Calories As A Function of Workout Duration") +theme_gray()
 
```    
*For the correlation heatmap, none of the variables have what is considered a strong relationship but interestingly the MaxHR and AvgHR have the greatest correlation of all the pairs that were looked at. The pair with the lowest was WoDuration and Calories. Moving on to the scatter plot, we see that there is a general appearance of a positive correlation between AvgHR and MaxHR but as noted with the correlation matrix, there is not truly a positive relationship here. Another thing to note is that the low fiber points are clustered more towards the lower part of the AvgHR when compared to the high fiber groups. Now for the last graph a big thing that jumped out at me is that there is quite a bit of spacing between the 60 minute and 76 minute mark. What this indicates that most of the workouts I did were around that time frame. Save for certain bars, the calories per minute are pretty consistent. I'd imagine it's because I was in a bigger caloric deficit when I did my longer workouts.*

5. Clustering
```{R}   
library(cluster)
    
    bigboi
    
 clust_data<-bigboi%>%dplyr::select(AvgHR,MaxHR,WoDuration,Calories) #Clustering on numeric Variables
 
 
 sil_width<-vector() #empty vector to hold mean sil width
for(i in 2:10){  
  kms <- kmeans(clust_data,centers=i) #compute k-means solution
  sil <- silhouette(kms$cluster,dist(clust_data)) #get sil widths
  sil_width[i]<-mean(sil[,3]) #take averages (higher is better)
}
ggplot()+geom_line(aes(x=1:10,y=sil_width))+scale_x_continuous(name="k",breaks=1:10)
#This graph shows that we should use two clusters as the silhouette width at that point is greatest! 
 
 
    set.seed(348)
pam2<-clust_data%>% scale %>% pam(k=3)
pam2 #setting Clustering

   
 yeetclust <-clust_data %>% mutate(cluster=as.factor(pam2$clustering))#saving so it can be plotted

 yeetclust %>% group_by(cluster) %>% summarise_if(is.numeric,mean,na.rm=T) #Means for clusters
 
 bigboi %>% slice(pam2$id.med)

 pam2$silinfo$avg.width
plot(pam2,which=2) 


final<-bigboi %>%  mutate(cluster=pam2$clustering)
 
library(GGally)
ggpairs(final,columns = 1:4, aes(color=as.factor(cluster)))
```
*For this section I started by clustering my data, and getting the silhouette widths in order to see what the best cluster number is! When I did this I saw that 3 clusters is the best and I went ahead and did the PAM for that. I also got the means for the clusters as well. Following this I got the goodness of fit, which gave me a value o 0.26 average silhouette width which indicates that the structure is weak. For the heart rate pair, we see that Clusters 1 and 2 are pretty grouped together and associated with high maximum heart rate and high average heart rate but that Cluster 3 is more distinct and associated with low values for both heart rate metrics. For the both the pairs of hear rate variables with Workout duration we do not see much variance of hear rate values except with the left side of the MaxHR and WoDuration which could be explained by a shorter more intense workout.*

<P style="page-break-before: always">
\newpage
```{r eval=F}
## paste this chunk into the ```{r setup} chunk at the top of your project 1 .Rmd file

knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = F, message = F,
tidy=TRUE, tidy.opts=list(width.cutoff=60), R.options=list(max.print=100))
```

...





